---
title: "Visualize experiments"
author: "Anna Boser"
date: "2022-11-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(data.table)
library(dplyr)
library(stringr)
library(ggplot2)
library(sf)
library(latex2exp)

source(here("file_paths.R"))
```

```{r}
experiment_name <- "fveg_water_buffer1000_2-10_gb"
experiment_path <- here("data", "4_for_analysis", "ML_outputs",  "experiments", experiment_name)
dataset <- "fveg"
months_ts <- TRUE
```

# ET over fallow lands vs predictions
```{r}
# read in the fallow fields predictions
fallow <- fread(here(experiment_path, "fallow.csv"))
fallow[fallow == -9999] <- NA # this is the na value

if (months_ts){
  # get a nice column of numeric months
  months <- select(fallow, names(fallow)[grepl("month", names(fallow))])
  fallow$month <- names(months)[max.col(months)] 
  fallow$month <- str_extract(fallow$month, '(?<=_)\\d+') %>% as.numeric()
}

if (years_ts){
  # get a nice column of numeric years
  years <- select(fallow, names(fallow)[grepl("year", names(fallow))])
  fallow$year <- names(years)[max.col(years)] 
  fallow$year <- str_extract(fallow$year, '(?<=_)\\d+') %>% as.numeric()
}
```

### A scatterplot with colors by month
```{r}
if (months_ts){
  # with different months
  r2 <- summary(lm(ET~ET_pred, data=fallow))$r.squared 
  Bias <- mean(fallow$ET_pred - fallow$ET, na.rm=TRUE)
  ggplot(fallow) +
    geom_jitter(aes(x=ET_pred, y=ET, color=as.factor(month)), alpha=0.2, size =.1) + 
    theme_classic() + 
    geom_abline(intercept=0,slope=1, color="red") + 
    # annotate("text", x=4, y=1.5, label= paste("R2:", round(r2, 3), "Bias:", round(Bias, 3))) + 
    xlab("Predicted natural ET (mm/month)") + 
    ylab("Observed fallow ET (mm/month)") +
    xlim(c(0,6*30)) + 
    ylim(c(0,6*30))
  }
print(paste("R2:", round(r2, 3), "Bias:", round(Bias, 3)))
```

### A scatterplot with months averaged out
```{r}
if (months_ts){
  # averaging over different months
  fallow_year <- fallow %>% group_by(x, y) %>% summarize(ET = mean(ET, na.rm=TRUE), 
                                                         ET_pred = mean(ET_pred, na.rm=TRUE))
} else {
  fallow_year <- fallow
}

r2 <- summary(lm(ET~ET_pred, data=fallow_year))$r.squared 
Bias <- mean(fallow_year$ET_pred - fallow_year$ET, na.rm=TRUE)
ggplot(fallow_year) +
  geom_jitter(aes(x=ET_pred, y=ET), alpha=0.2, size =.1) + 
  theme_classic() + 
  geom_abline(intercept=0,slope=1, color="red") + 
  # annotate("text", x=3, y=1.5, label= paste("R2:", round(r2, 3), "Bias:", round(Bias, 3))) + 
  xlab("Predicted natural ET (mm/month)") + 
  ylab("Observed fallow ET (mm/month)") +
    xlim(c(0,4*30)) + 
    ylim(c(0,4*30))

print(paste("R2:", round(r2, 3), "Bias:", round(Bias, 3)))
```

### A time series 
```{r}
if (months_ts){
  # plot out the time series
  fallow_ts <- fallow %>% group_by(month) %>% summarize(ET = mean(ET, na.rm=TRUE), 
                                                         ET_pred = mean(ET_pred, na.rm=TRUE))
  
  ggplot(fallow_ts) + 
    geom_line(aes(x=month, y=ET), color="red") + 
    geom_line(aes(x=month, y=ET_pred), color="black") 
}
```

Here I just got curious about what ag ET vs fallow ET look like. Seems like fallow ET is probably artificially high. 
```{r}
# ag <- fread(here("data", "3_for_counterfactual", "agriculture", "agriculture.csv"))
# ag$month <- str_extract(ag$month, '(?<=_)\\d+') %>% as.numeric()
# 
# # plot out the time series
# ag_ts <- ag %>% group_by(month) %>% summarize(ET = mean(ET, na.rm=TRUE))
# 
# ggplot(ag_ts) + 
#   geom_line(aes(x=month, y=ET), color="red") + 
#   geom_line(data=fallow_ts, aes(x=month, y=ET), color="black")
```


# Spatial CV

```{r}
# read in the spatial_CV predictions
cv <- fread(here(experiment_path, "crossval_predictions_test.csv"))
cv[cv == -9999] <- NA # this is the na value

if (months_ts){
  # get a nice column of numeric months
  months <- select(cv, names(cv)[grepl("month", names(cv))])
  cv$month <- names(months)[max.col(months)] 
  cv$month <- str_extract(cv$month, '(?<=_)\\d+') %>% as.numeric()
}

cv$mean_dist <- cv$fold_size/6 # the average distance between a held out point and the border to the hold out
```

## map out the performance of the model as hold out sizes increase

```{r}
if (months_ts){
  gridded_cv_metrics <- cv %>% 
    group_by(x, y, cv_fold, fold_size) %>%
    summarize(ET = mean(ET, na.rm = TRUE), 
              ET_pred = mean(ET_pred, na.rm = TRUE)) %>%
    group_by(cv_fold, fold_size) %>%
    summarize(r2 = 1 - mean((ET_pred - ET)^2)/mean((ET - mean(ET))^2), 
              Bias = mean(ET_pred - ET, na.rm = TRUE), 
              RMSE = sqrt(mean(ET_pred - ET, na.rm = TRUE)^2), 
              mae = mean(abs((ET_pred - ET)), na.rm = TRUE),
              ET = mean(ET, na.rm = TRUE), 
              ET_pred = mean(ET_pred, na.rm = TRUE))
} else {
  gridded_cv_metrics <- cv %>%
    group_by(cv_fold, fold_size) %>%
    summarize(r2 = 1 - mean((ET_pred - ET)^2)/mean((ET - mean(ET))^2),
              Bias = mean(ET_pred - ET, na.rm = TRUE),
              RMSE = sqrt(mean(ET_pred - ET, na.rm = TRUE)^2),
              mae = mean(abs((ET_pred - ET)), na.rm = TRUE),
              ET = mean(ET, na.rm = TRUE),
              ET_pred = mean(ET_pred, na.rm = TRUE))
}
```

Turn the cv fold back into something meaningful
```{r}
# first turn "fold" into x and y coordinates
gridded_cv_metrics$x <- as.numeric(str_extract(gridded_cv_metrics$cv_fold, "[-.0-9]*(?=,)")) + gridded_cv_metrics$fold_size/200000
gridded_cv_metrics$y <- as.numeric(str_extract(gridded_cv_metrics$cv_fold, "(?<=,)[-.0-9]*")) + gridded_cv_metrics$fold_size/200000
```


```{r}
# study area and counties for plotting
study_area <- st_read(study_area_loc) %>% st_transform(st_crs("+proj=longlat +datum=WGS84"))
counties <- st_read(counties_loc) %>% filter(STATEFP == "06") %>% st_transform(st_crs("+proj=longlat +datum=WGS84"))

gridded_cv_metrics$RMSE <- ifelse(gridded_cv_metrics$RMSE>1.5, 1.5, gridded_cv_metrics$RMSE)
gridded_cv_metrics$Bias <- ifelse(gridded_cv_metrics$Bias>1, 1, gridded_cv_metrics$Bias)
gridded_cv_metrics$Bias <- ifelse(gridded_cv_metrics$Bias<(-1), -1, gridded_cv_metrics$Bias)

filter(gridded_cv_metrics) %>%
  ggplot() + 
  geom_sf(data = counties, color=alpha("white",1), size = .2) + 
  geom_raster(aes(x=x, y=y, fill=Bias)) +
  facet_wrap(vars(fold_size/1000)) +
  scale_fill_distiller(palette="Spectral", direction = -1, limits = c(-1*30,1*30)) +
  geom_sf(data = study_area, fill=alpha("red",0), color = "black", size = .2) + 
  theme_classic() +
  xlim(c(-122.92, -118)) + 
  ylim(c(34.7, 40.754)) + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

filter(gridded_cv_metrics) %>%
  ggplot() + 
  geom_sf(data = counties, color=alpha("white",1), size = .2) + 
  geom_raster(aes(x=x, y=y, fill=RMSE)) +
  facet_wrap(vars(fold_size/1000)) +
  scale_fill_distiller(palette="Spectral", direction = -1, limits = c(0,1.5*30)) +
  geom_sf(data = study_area, fill=alpha("red",0), color = "black", size = .2) + 
  theme_classic() +
  xlim(c(-122.92, -118)) + 
  ylim(c(34.7, 40.754)) + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

## Plot out the performance of the model as hold out sizes get bigger relative to the distance between ag and natural training data

First get the distribution of distances between ag and natural data (specifically the test data)
```{r}
dist_filename <- here(distance_distribution_path, paste0(dataset, ".csv"))

if (file.exists(dist_filename)){
  library(parallel)
  library(tictoc)
  
  dist <- fread(dist_filename)
} else {
  stop('No distance file for this dataset. Make one using 3_analysis/1_additional_data_manipulation/2_ag_natural_distance.R')
}
```



```{r}
if (months_ts){
  cv_metrics <- cv %>% 
    group_by(x, y, fold_size, mean_dist) %>%
    summarize(ET = mean(ET, na.rm = TRUE), 
              ET_pred = mean(ET_pred, na.rm = TRUE)) %>%
    group_by(fold_size, mean_dist) %>%
    summarize(r2 = 1 - mean((ET_pred - ET)^2)/mean((ET - mean(ET))^2), 
              Bias = mean(ET_pred - ET, na.rm = TRUE), 
              RMSE = sqrt(mean(ET_pred - ET, na.rm = TRUE)^2), 
              mae = mean(abs((ET_pred - ET)), na.rm = TRUE),
              ET = mean(ET, na.rm = TRUE), 
              ET_pred = mean(ET_pred, na.rm = TRUE))
} else {
  cv_metrics <- cv %>%
    group_by(fold_size, mean_dist) %>%
    summarize(r2 = 1 - mean((ET_pred - ET)^2)/mean((ET - mean(ET))^2),
              Bias = mean(ET_pred - ET, na.rm = TRUE),
              RMSE = sqrt(mean(ET_pred - ET, na.rm = TRUE)^2),
              mae = mean(abs((ET_pred - ET)), na.rm = TRUE),
              ET = mean(ET, na.rm = TRUE),
              ET_pred = mean(ET_pred, na.rm = TRUE))
}
```


```{r}
coef = 2/5

ggplot(NULL) + 
  stat_density(data = dist, aes(x=mindist/1000), fill = "grey10", alpha = .3) + 
  geom_line(data = cv_metrics, aes(x=mean_dist/1000, y = r2*coef), color = "red") +
  geom_point(data = cv_metrics, aes(x=mean_dist/1000, y = r2*coef), color = "red") +
  xlab(TeX("Distance of agricultural pixel to nearest natural pixel (km) \n Average distance of point to border of hold-out (km)")) + 
  theme_bw() + 
  theme(axis.title.x = element_text(vjust=-2.5)) +
  xlim(c(0, max(cv_metrics$mean_dist/1000))) + 
  scale_y_continuous(
    name = "Distribution of agricultural pixels",
    sec.axis = sec_axis(~./coef, name=TeX("Model performance ($R^{2}$)")))
```

### Plot the performance for the different months too
```{r}
if (months_ts){
  cv_metrics <- cv %>% 
    group_by(fold_size, month, mean_dist) %>%
    summarize(r2 = 1 - mean((ET_pred - ET)^2)/mean((ET - mean(ET))^2), 
              Bias = mean(ET_pred - ET, na.rm = TRUE), 
              RMSE = sqrt(mean(ET_pred - ET, na.rm = TRUE)^2), 
              mae = mean(abs((ET_pred - ET)), na.rm = TRUE),
              ET = mean(ET, na.rm = TRUE), 
              ET_pred = mean(ET_pred, na.rm = TRUE))
  
  coef = 2/5
  
  ggplot(NULL) + 
    stat_density(data = dist, aes(x=mindist/1000), fill = "grey10", alpha = .3) + 
    geom_line(data = cv_metrics, aes(x=mean_dist/1000, y = r2*coef, color = as.factor(month))) +
    geom_point(data = cv_metrics, aes(x=mean_dist/1000, y = r2*coef, color = as.factor(month))) +
    xlab(TeX("Distance of agricultural pixel to nearest natural pixel (km) \n Average distance of point to border of hold-out (km)")) + 
    theme_bw() + 
    theme(axis.title.x = element_text(vjust=-2.5)) +
    xlim(c(0, max(cv_metrics$mean_dist/1000))) + 
    scale_y_continuous(
      name = "Distribution of agricultural pixels",
      sec.axis = sec_axis(~./coef, name=TeX("Model performance ($R^{2}$)")))
}
```

## Some scatterplots and time series

Make these using the CV with the fold size that is closest to the double of the mean distance between ag and natural (a bit conservative but only slightly)
```{r}
mean_dist <- mean(dist$mindist)
folds <- unique(cv$mean_dist)
best_fold <- folds[abs(folds-(mean_dist)) == min(abs(folds-(mean_dist)))]
cv_best_fold <- filter(cv, mean_dist == best_fold)
print(cv_best_fold$fold_size %>% unique())
```

### A scatterplot with colors by month
```{r}
if (months_ts){
  # with different months
  r2 <- summary(lm(ET~ET_pred, data=cv_best_fold))$r.squared 
  Bias <- mean(cv_best_fold$ET_pred - cv_best_fold$ET, na.rm=TRUE)
  ggplot(cv_best_fold) +
    geom_jitter(aes(x=ET_pred, y=ET, color=as.factor(month)), alpha=0.2, size =.1) + 
    theme_classic() + 
    geom_abline(intercept=0,slope=1, color="red") + 
    # annotate("text", x=4.5, y=1.5, label= paste("R2:", round(r2, 3), "Bias:", round(Bias, 3))) + 
    xlab("Predicted natural ET (mm/month)") + 
    ylab("Observed fallow ET (mm/month)") +
    xlim(c(0,8*30)) + 
    ylim(c(0,8*30))
}

print(paste("R2:", round(r2, 3), "Bias:", round(Bias, 3)))
```

### A scatterplot with months averaged out
```{r}
if (months_ts){
  # averaging over different months
  cv_best_fold_year <- cv_best_fold %>% group_by(x, y) %>% summarize(ET = mean(ET, na.rm=TRUE), 
                                                         ET_pred = mean(ET_pred, na.rm=TRUE))
} else {
  cv_best_fold_year <- cv_best_fold
}

r2 <- summary(lm(ET~ET_pred, data=cv_best_fold_year))$r.squared 
Bias <- mean(cv_best_fold_year$ET_pred - cv_best_fold_year$ET, na.rm=TRUE)
ggplot(cv_best_fold_year) +
  geom_jitter(aes(x=ET_pred, y=ET), alpha=0.02, size =.01) + 
  theme_classic() + 
  geom_abline(intercept=0,slope=1, color="red") + 
  # annotate("text", x=4.5, y=1.5, label= paste("R2:", round(r2, 3), "Bias:", round(Bias, 3))) + 
  xlab("Predicted natural ET (mm/month)") + 
  ylab("Observed fallow ET (mm/month)") +
    xlim(c(0,6*30)) + 
    ylim(c(0,6*30))

print(paste("R2:", round(r2, 3), "Bias:", round(Bias, 3)))
```

### A time series 
```{r}
if (months_ts){
  # plot out the time series
  cv_best_fold_ts <- cv_best_fold %>% group_by(month) %>% summarize(ET = mean(ET, na.rm=TRUE), 
                                                         ET_pred = mean(ET_pred, na.rm=TRUE))
  
  ggplot(cv_best_fold_ts) + 
    geom_line(aes(x=month, y=ET), color="red") + 
    geom_line(aes(x=month, y=ET_pred), color="black") + 
    geom_line(data = fallow_ts, aes(x=month, y=ET), color="blue") + 
    geom_line(data = fallow_ts, aes(x=month, y=ET_pred), color="green") 
}
```
