---
title: "irrigation comparison"
author: "Anna Boser"
date: '2022-05-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(stringr)
library(sf)
library(tmap)
library(data.table)
library(tidyr)
library(dplyr)
library(latex2exp)
library(ggplot2)
library(lfe)
library(rgdal)
library(ggnewscale) # allows you to plot multiple things with different scales on one ggplot

source(here("file_paths.R"))
```


## Read in the data
```{r}
experiment_name <- "fveg_11-30_gb_test_set"
experiment_path <- here("data", "4_for_analysis", "ML_outputs",  "experiments", experiment_name)
months_ts <- TRUE
```

```{r}
data <- fread(file = here(experiment_path, "agriculture.csv"))

data[data == -9999] <- NA # this is the na value

if (months_ts){
  # get a nice column of numeric months
  months <- select(data, names(data)[grepl("month", names(data))])
  data$month <- names(months)[max.col(months)] 
  data$month <- str_extract(data$month, '(?<=_)\\d+') %>% as.numeric()
}
```

## average over months
```{r}
if (months_ts == TRUE){
  # Average over monthgroups
  data <- data %>% group_by(x, y, Soil, TWI, Elevation, Aspect, Slope) %>% summarize(ET = mean(ET, na.rm=TRUE), 
                                                  ET_pred = mean(ET_pred, na.rm=TRUE), 
                                                  ag_ET = ET-ET_pred, 
                                                PET = mean(PET, na.rm=TRUE))
}
```

# add county and crop type information 
```{r}
data$x <- round(data$x, 7)
data$y <- round(data$y, 7)

# add counties
counties <- fread(county_table_loc)
counties$cell <- NULL
counties$x <- round(counties$x, 7)
counties$y <- round(counties$y, 7)
data <- merge(data, counties, by = c("x", "y"), all.x=TRUE)

# add_crops
crops <- fread(crops_table_loc)
crops$cell <- NULL
crops$x <- round(crops$x, 7)
crops$y <- round(crops$y, 7)

data <- merge(data, crops, by = c("x", "y"), all.x=TRUE)
data$cropnames <- ifelse(data$cropnames %in% c("Unclassified fallow", "Idle"), "Fallow or idle", data$cropnames) # rename unclassified fallow to just fallow

# filter out urban
data <- filter(data, cropnames != "Urban - residential, commercial, and industrial, unsegregated")

# remove fallow and grain crops since these are unlikely to be irrigated
data <- filter(data, cropnames != "Fallow or idle")
data <- filter(data, cropnames != "Grain and hay crops")
```

Assign a cluster to get the proper standard errors
```{r}
# create clusters based on location. Function creates clusters of size dist km. 
assign_cluster <- function(x, y, dist){
  
  x_size = dist/89 # 1 degree lon (x) = 89km = 89000m
  y_size = dist/111 # 1 degree lat (y) = 111km = 111000m
  
  x_fold = floor(x/x_size)*x_size
  y_fold = floor(y/y_size)*y_size
  
  cv_fold = paste(x_fold, y_fold, sep = ",")
  
  return(cv_fold)
}

data$cluster20 <- mapply(assign_cluster, data$x, data$y, 20)
```

I then need to get the total water ETd, so I multiply by the number of days in my observation window: 183
```{r}
data$ag_ET_year <- data$ag_ET*183
data$ET_year <- data$ET*183
```

Use a clustered standard errors regression to retrieve the water consumption by county
```{r}
# get confidence intervals using clustered standard errors
ag_ET_20 <- felm(ag_ET_year~NAME-1 | 0 | 0 | cluster20, data)

# extract the mean and ci values from each
df_from_felm <- function(input_lm){
  ET_df <- as.data.frame(cbind(input_lm$coefficients, input_lm$cse))
  colnames(ET_df) <- c("ag_ET_mm_year", "ag_ET_se")
  ET_df$ag_ET_min <- ET_df$ag_ET_mm_year-(2*ET_df$ag_ET_se)
  ET_df$ag_ET_max <- ET_df$ag_ET_mm_year+(2*ET_df$ag_ET_se)
  ET_df$NAME <- substring(rownames(ET_df), 5)
  return(ET_df)
}

ag_ET_df <- df_from_felm(ag_ET_20)
```

```{r}
# I'm also interested in knowing what the irrigation efficiency is if I totally ignore natural ET -- so on the total ET

# get confidence intervals using clustered standard errors
ET_20 <- felm(ET_year~NAME-1 | 0 | 0 | cluster20, data)

# extract the mean and ci values from each
df_from_felm <- function(input_lm){
  ET_df <- as.data.frame(cbind(input_lm$coefficients, input_lm$cse))
  colnames(ET_df) <- c("ET_mm_year", "ET_se")
  ET_df$ET_min <- ET_df$ET_mm_year-(2*ET_df$ET_se)
  ET_df$ET_max <- ET_df$ET_mm_year+(2*ET_df$ET_se)
  ET_df$NAME <- substring(rownames(ET_df), 5)
  return(ET_df)
}

ET_df <- df_from_felm(ET_20)
```

```{r}
# Merge the two together
ET_df <- merge(ET_df, ag_ET_df, by = "NAME")
```

Add in some more information about the counties
```{r}
# average over county
county_avg <- data %>%
  group_by(NAME) %>% # group by county here
  summarize(
    lon = mean(x, na.rm=TRUE),
    lat = mean(y, na.rm=TRUE),
    soil = mean(Soil, na.rm=TRUE),
    ET = mean(ET, na.rm=TRUE), # all these are mm/day
    ET_pred = mean(ET_pred, na.rm=TRUE),
    ag_ET = mean(ag_ET, na.rm=TRUE),
    PET = mean(PET, na.rm=TRUE),
    n = n()) #get the number of pixels in each county

# add other information about the counties back in 
ET_df <- merge(ET_df, county_avg, by = "NAME")
```


## USGS Irrigation data

### Read in USGS data. Make a single irrigation dataset with 2010 and 2015 data
```{r}
irrigation2015 <- read.csv(here(usgs_irr_path,
                       "2015",
                       "water_use"), sep = "\t")[-1,] #first row just tells you the size of the entries; remove

irrigation2010 <- read.csv(here(usgs_irr_path,
                       "2010",
                       "water_use"), sep = "\t")[-1,] #first row just tells you the size of the entries; remove

irrigation2015$year <- 2015
irrigation2010$year <- 2010

irrigation <- rbind(irrigation2015, irrigation2010)

rm(irrigation2010)
rm(irrigation2015)
```

### Clean USGS irrigation data
Change units to kg/day and meters squared
```{r}
irrigation$NAME <- str_remove(irrigation$county_nm, " County") # remove the "County" suffix to each county name in order to match to county shapefile names


# convert billions of gallons per day to kg per day
conversion <- function(Mgal.d){
  Mgal.d = as.numeric(Mgal.d)
  kg.d = Mgal.d * 3.785411784 * 1000000
}

to_convert = names(irrigation)[c(6:10, 15)] #the names of the columns that are in Mgal/d and need to be converted
irrigation <- mutate(irrigation, across(to_convert, conversion))


# convert thousands of acres to square meters
conversion <- function(tacre){
  tacre = as.numeric(tacre)
  m2 = tacre * 1000 * 4046.86 # conversion to acres and conversion to square meters
}

to_convert = names(irrigation)[c(11:14)] #the names of the columns that are in thousand of acres and need to be converted

irrigation <- mutate(irrigation, across(to_convert, conversion))

# add a mm/day and mm/year variable
# keep in mind the column names no longer reflect the true units
irrigation$mm.day <- irrigation$Irrigation..Crop.total.self.supplied.withdrawals.for.crops..fresh..in.Mgal.d/irrigation$Irrigation..Crop.total.irrigation.for.crops..in.thousand.acres

irrigation$mm.year <- irrigation$mm.day*365

# also get the percent drip and flood irrigation for each county
irrigation$drip <- irrigation$Irrigation..Crop.microirrigation.for.crops..in.thousand.acres/irrigation$Irrigation..Crop.total.irrigation.for.crops..in.thousand.acres
irrigation$flood <- irrigation$Irrigation..Crop.surface.irrigation.for.crops..in.thousand.acres/irrigation$Irrigation..Crop.total.irrigation.for.crops..in.thousand.acres
```


```{r}
# take the average irrigation over different years to simplify
irrigation_avg <- irrigation %>% 
  group_by(NAME) %>% 
  summarize(mm.year = mean(mm.year),
            mm.day = mean(mm.day),
            flood = mean(flood), 
            drip = mean(drip))
```

```{r}
# merge ET and irrigation datasets to get the irrigation efficiency
efficiency_df <- base::merge(irrigation_avg, ET_df, by="NAME", all.x = TRUE)
efficiency_df$irrigation_efficiency <- efficiency_df$ag_ET_mm_year*100/efficiency_df$mm.year
efficiency_df$irrigation_efficiency_min <- efficiency_df$ag_ET_min*100/efficiency_df$mm.year
efficiency_df$irrigation_efficiency_max <- efficiency_df$ag_ET_max*100/efficiency_df$mm.year
efficiency_df$prop_total_ET <- efficiency_df$ET_mm_year*100/efficiency_df$mm.year
efficiency_df$prop_total_ET_min <- efficiency_df$ET_min*100/efficiency_df$mm.year
efficiency_df$prop_total_ET_max <- efficiency_df$ET_max*100/efficiency_df$mm.year
```

## Do some plotting!

### Read in county shapefile and merge dfs
```{r}
# study area and counties
study_area <- st_read(study_area_loc) %>% st_transform(st_crs("+proj=longlat +datum=WGS84"))
counties <- st_read(counties_loc) %>% filter(STATEFP == "06") %>% st_transform(st_crs("+proj=longlat +datum=WGS84"))
counties_irrigation <- base::merge(counties, efficiency_df, by = "NAME")
```


### Plot 1: irrigation use and agricultural water consumption
```{r}
ggplot() + 
  geom_sf(data = counties_irrigation, aes(fill = mm.year), color=alpha("red",0)) + 
  scale_fill_gradient2(name="Irrigation (mm/year)", low = "red", mid = "white", high = "deepskyblue4") + 
  # geom_sf(data = counties_irrigation, fill=alpha("red",0), color=alpha("grey64")) + 
  geom_sf(data = study_area, fill=alpha("red",0), color = "black", size = .2) + 
  new_scale_fill() +
  geom_raster(data = data[sample(nrow(data), size = nrow(data), replace = FALSE),], aes(x=x, y=y, fill=ag_ET_year), alpha=0.75) +
  scale_fill_gradientn(name="Agricultural water consumption (mm/year)", colours = c("darkgoldenrod4", "darkgoldenrod2", "khaki1", "lightgreen", "turquoise3", "deepskyblue3", "mediumblue", "navyblue", "midnightblue", "black"), limits = c(-2.5*183, 8*183)) +
  theme_void() + 
  theme(legend.position = c(.8, .75), 
        legend.direction = "horizontal", 
        legend.title = element_blank(), 
        legend.margin = margin(25))
```

### Plot 1b: irrigation use and ET
```{r}
ggplot() + 
  geom_sf(data = counties_irrigation, aes(fill = mm.year), color=alpha("red",0)) + 
  scale_fill_gradient2(name="Irrigation (mm/year)", low = "red", mid = "white", high = "deepskyblue4") + 
  # geom_sf(data = counties_irrigation, fill=alpha("red",0), color=alpha("grey64")) + 
  geom_sf(data = study_area, fill=alpha("red",0), color = "black", size = .2) + 
  new_scale_fill() +
  geom_raster(data = data[sample(nrow(data), size = nrow(data), replace = FALSE),], aes(x=x, y=y, fill=ET_year), alpha=0.75) +
  scale_fill_gradientn(name="Observed ET (mm/year)", colours = c("darkgoldenrod4", "darkgoldenrod2", "khaki1", "lightgreen", "turquoise3", "deepskyblue3", "mediumblue", "navyblue", "midnightblue", "black"), limits = c(-2.5*183, 8*183)) +
  theme_void() + 
  theme(legend.position = c(.8, .75), 
        legend.direction = "horizontal", 
        legend.title = element_blank(), 
        legend.margin = margin(25))
```

### Plot 2: irrigation use and agricultural water consumption in the central valley
```{r}
# remove counties that don't have any CV agriculture
cv_data <- counties_irrigation
cv_data$irrigation_efficiency <- ifelse(counties_irrigation$n > 50000, cv_data$irrigation_efficiency, NA)

ggplot() + 
  geom_sf(data = cv_data, aes(fill = irrigation_efficiency), color=alpha("red",0)) + 
  geom_text(data = cv_data, aes(x = lon, y = lat, label = round(irrigation_efficiency, 0)), size = 2) + 
  scale_fill_gradient2(name = "Irrigation efficiency (%)", low = "red", mid = "white", high = "#70AD47", limits = c(0, 25)) + 
  geom_sf(data = study_area, fill=alpha("red",0), color = "black", size = .2) + 
  theme_void() + 
  theme(legend.position = c(.85, .77)) 
```

### Plot 2b: total ET/diverted irrigation
```{r}
# remove counties that don't have any CV agriculture
cv_data <- counties_irrigation
cv_data$prop_total_ET <- ifelse(counties_irrigation$n > 50000, cv_data$prop_total_ET, NA)

ggplot() + 
  geom_sf(data = cv_data, aes(fill = prop_total_ET), color=alpha("red",0)) + 
  geom_text(data = cv_data, aes(x = lon, y = lat, label = round(prop_total_ET, 0)), size = 2) + 
  scale_fill_gradient2(name = "Observed ET/Irrigation (%)", low = "red", mid = "white", high = "#70AD47", limits = c(0, 110)) + 
  geom_sf(data = study_area, fill=alpha("red",0), color = "black", size = .2) + 
  theme_void() + 
  theme(legend.position = c(.85, .77)) 
```
```{r}
# average drip irrigated by county
mean(filter(counties_irrigation, n > 50000)$drip, na.rm = TRUE)
mean(filter(counties_irrigation, n > 50000)$flood, na.rm = TRUE)
```

## Statsitics
```{r}
# however, in order to get confidence intervals I will need the entire dataset
data_irr <- merge(data, irrigation_avg, by = "NAME", all.x = TRUE, all.y = FALSE)
data_irr$irrigation_efficiency <- data_irr$ag_ET_year*100/data_irr$mm.year
data_irr$prop_total_ET <- data_irr$ET_year*100/data_irr$mm.year

# get confidence intervals using clustered standard errors
total_eff <- felm(irrigation_efficiency~1 | 0 | 0 | cluster20, data_irr)
total_prop <- felm(prop_total_ET~1 | 0 | 0 | cluster20, data_irr)
```

```{r}
# extract the mean and ci values from each
df_from_felm <- function(input_lm){
  ET_df <- as.data.frame(cbind(input_lm$coefficients, input_lm$cse))
  colnames(ET_df) <- c("mean", "se")
  ET_df$min <- ET_df$mean-(2*ET_df$se)
  ET_df$max <- ET_df$mean+(2*ET_df$se)
  ET_df$NAME <- substring(rownames(ET_df), 5)
  return(ET_df)
}

total_eff_df <- df_from_felm(total_eff)

total_prop_df <- df_from_felm(total_prop)

print("total_eff_df")
print(total_eff_df)
print("total_prop_df")
print(total_prop_df)
```

### explaining trends in irrigation efficiency
```{r}
irr_df <- filter(counties_irrigation, n > 50000)
summary(lm(irrigation_efficiency~flood, data = irr_df))
summary(lm(irrigation_efficiency~drip, data = irr_df))
summary(lm(irrigation_efficiency~lon, data = irr_df))
summary(lm(irrigation_efficiency~soil, data = irr_df))
summary(lm(irrigation_efficiency~PET, data = irr_df))
summary(lm(irrigation_efficiency~lat, data = irr_df))
summary(lm(irrigation_efficiency~drip+flood, data = irr_df))
```

```{r}
counties_irrigation %>% filter(n > 50000) %>% select(NAME, irrigation_efficiency, irrigation_efficiency_min, irrigation_efficiency_max) %>% as.data.frame()
```

