---
title: "orchard age analysis"
author: "Anna Boser"
date: "2023-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(here)
library(stringr)
library(sf)
library(data.table)
library(tidyr)
library(dplyr)
library(ggplot2)
library(latex2exp)
library(lfe)

source(here("file_paths.R"))
source(here("helper_functions.R"))
```

# read in data; only keep YP or orchards (vineyards; citrus & subtropical; deciduous fruits & nuts)
```{r}
experiment_name <- "fallow0.05,2_4-18_gb"
experiment_path <- here("data", "4_for_analysis", "ML_outputs",  "experiments", experiment_name)
data <- fread(file = here(experiment_path, "agriculture_yearly.csv"))# , nrows = 100000)
# test <- fread(here(test_data_path, paste0("fallow0,2_test.csv"))) # the pixels not used in training or validation 
test <- fread(here(test_data_path, paste0("fallow0.05,2_test.csv"))) # the pixels not used in training or validation 

report <- here(experiment_path, "8_report.txt")
if (file.exists(report)) {file.remove(report)} # start fresh every time you render
```

## clean data
```{r}
data <- filter(data, coverage_fraction>.5)
data <- filter(data, cropnames %in% c("Young Perennial", "Deciduous fruits and nuts", "Vineyards", "Citrus and subtropical"))
```

```{r}
# create clusters based on location and year. Function creates clusters of size dist km. 
data$cluster <- mapply(assign_cluster, data$x, data$y, 75)

# scale to cm/year instead of mm/month
# scalar = 1.2 # turns mm/month to cm/year
# data <- scale(data, scalar)

# control for variation across years
data <- year_control(data)
```

# load in the crop type data for 2014-2020
```{r}
get_crop <- function(year){
  crop <- fread(here(crops_dwr_table_path, paste0("crops_all_dwr_fallow", year, ".csv")))
  crop$year <- year
  return(crop)
}
crop_data <- rbindlist(lapply(c(2014, 2016, 2018, 2019, 2020), get_crop), fill = TRUE)
```



# If orchard: Add a column indicating how far past YP they are and a helper column
```{r}
data$year <- as.numeric(as.character(data$year)) # was factor -- first turn to character to get rid of levels and then numeric
data$loc <- paste(sprintf("%.7f", data$x), sprintf("%.7f", data$y))
orchards <- filter(data, cropnames != "Young Perennial") # this is what you will use to make your figure

crop_data$year <- as.numeric(as.character(crop_data$year)) # was factor -- first turn to character to get rid of levels and then numeric
crop_data$loc <- paste(sprintf("%.7f", crop_data$x), sprintf("%.7f", crop_data$y))

# maximum year YP was seen
max_YP_year <- crop_data %>%
  filter(cropnames == "Young Perennial") %>%
  group_by(loc) %>%
  mutate(max_YP_year = max(year)) %>%
  filter(max_YP_year == year) %>% # only one observation per pixel
  select(loc, max_YP_year) %>%
  ungroup()

max_YP_year <- max_YP_year[!duplicated(paste(max_YP_year$loc, max_YP_year$max_YP_year)),]

orchards <- merge(orchards, max_YP_year, by = c("loc"), all.x = TRUE, all.y = FALSE)

# first year that this pixel is not an orchard -- does not necessarily need to be YP since it could be replaced by anything
min_YP_year <- crop_data %>%
  filter(!(cropnames %in% c("Young Perennial", "Deciduous fruits and nuts", "Vineyards", "Citrus and subtropical"))) %>%
  group_by(loc) %>%
  mutate(min_YP_year = min(year)) %>%
  filter(min_YP_year == year) %>% # only one observation per pixel
  select(loc, min_YP_year) %>%
  ungroup()

min_YP_year <- min_YP_year[!duplicated(paste(min_YP_year$loc, min_YP_year$min_YP_year)),]

orchards <- merge(orchards, min_YP_year, by = c("loc"), all.x = TRUE, all.y = FALSE)

```

```{r}
# how many years post YP
orchards$age <- orchards$year - orchards$max_YP_year
# max_age <- max(orchards$age, na.rm = TRUE)
orchards$age <- ifelse(is.na(orchards$age) | orchards$age < 0, "Intermediate", orchards$age)
# orchards$age <- ifelse(orchards$age < 0, orchards$age + 2*(max_age + 1), orchards$age)
# orchards$age <- as.factor(orchards$age)

# how close to being ripped out (t minus)
orchards$old <- orchards$year - orchards$min_YP_year

orchards$age <- ifelse(is.na(orchards$old) | orchards$old > 0, orchards$age, paste0(" ", orchards$old))

unique(orchards$age)
```

# If YP: Add a column indicating how far from becoming an orchard YP fields are are

```{r}
min_orch_year <- crop_data %>%
  filter(cropnames != "Young Perennial") %>%
  group_by(loc) %>%
  mutate(min_orch_year = min(year)) %>%
  filter(min_orch_year == year) %>% # only one observation per pixel
  select(loc, min_orch_year) %>%
  ungroup()

min_orch_year <- min_orch_year[!duplicated(paste(min_orch_year$loc, min_orch_year$min_orch_year)),]

YP <- filter(data, cropnames == "Young Perennial")
YP <- merge(YP, min_orch_year, by = c("loc"), all.x = TRUE, all.y = FALSE)
```

```{r}
YP$age <- YP$year - YP$min_orch_year
YP <- filter(YP, YP$age < 0) 
YP$age <- YP$age

unique(YP$age)

```


# Plot out ag ET as a function of both of these. 
```{r}
age_df <- rbind(orchards, YP, fill=TRUE)
age_df$age <- factor(age_df$age)

ET_c <- felm(ag_ET~age-1 | 0  | 0 | cluster, age_df) # -1 means no intercept
ET_df <- df_from_felm(ET_c, name_length = 4, var="age")
rm(ET_c)
```

```{r}
# Filter out the "Intermediate age" category and store the values
all_others_data <- ET_df[ET_df$age == "Intermediate", ]

ET_df$category <- factor(
  case_when(
    grepl("^Int", ET_df$age) ~ "Intermediate age",
    grepl("^-", ET_df$age) ~ "Years to bearing fruit",
    grepl("^ -", ET_df$age) ~ "Years to removal",
    TRUE ~ "Years bearing fruit"
  ), levels = c("Years to bearing fruit", "Years bearing fruit", "Intermediate age", "Years to removal")
)

ggplot(ET_df) + 
  # Color the points by category
  geom_pointrange(aes(x = factor(age, levels = c("-4", "-3", "-2", "-1", "1", "2", "3", "4", "5", "Intermediate", " -4", " -3", " -2", " -1")), y = ET, ymin = min, ymax = max, color = category)) +
  
  # Add a horizontal line for the "Intermediate age" mean
  geom_hline(aes(yintercept = all_others_data$ET), color = "grey50", linetype="dashed") +
  
  # Add confidence interval band
  geom_rect(aes(ymin = all_others_data$min, ymax = all_others_data$max, xmin = 0, xmax = length(unique(ET_df$age)) + 1), fill = "grey50", alpha = 0.02) +
  
  # Manual legend for the categories and "intermediate age"
  scale_color_manual(name = "", values = c("Years to bearing fruit" = "palegreen2", "Years bearing fruit" = "green3", "Years to removal" = "darkgreen", "Intermediate age" = "grey50")) +
  
  # Axis labels
  xlab("Orchard age") +
  ylab("Agricultural ET (mm/month)") + 
  theme_bw() + 
  theme(legend.position = c(.8, .25), 
        axis.text.x = element_text(angle = 20, hjust = 1))

ggsave(here(experiment_path, paste0("Figure_S_Orchard_age.pdf")), 
       width = 17.8,
       height = 11.1,
       units = "cm")
```


# Plot out ag ET as a function of both of these. 

```{r}

for (crop in c("Deciduous fruits and nuts", "Vineyards", "Citrus and subtropical")){
  
  age_df_filt <- filter(age_df, cropnames %in% c(crop, "Young Perennial"))

  ET_c <- felm(ag_ET~age-1 | 0  | 0 | cluster, age_df_filt) # -1 means no intercept
  ET_df <- df_from_felm(ET_c, name_length = 4, var="age")
  rm(ET_c)
  
  # Filter out the "Intermediate age" category and store the values
  all_others_data <- ET_df[ET_df$age == "Intermediate", ]
  
  ET_df$category <- factor(
    case_when(
      grepl("^Int", ET_df$age) ~ "Intermediate age",
      grepl("^-", ET_df$age) ~ "Years to bearing fruit",
      grepl("^ -", ET_df$age) ~ "Years to removal",
      TRUE ~ "Years bearing fruit"
    ), levels = c("Years to bearing fruit", "Years bearing fruit", "Intermediate age", "Years to removal")
  )
  
  plot <- ggplot(ET_df) + 
    # Color the points by category
    geom_pointrange(aes(x = factor(age, levels = c("-4", "-3", "-2", "-1", "1", "2", "3", "4", "5", "Intermediate", " -4", " -3", " -2", " -1")), y = ET, ymin = min, ymax = max, color = category)) +
    
    # Add a horizontal line for the "Intermediate age" mean
    geom_hline(aes(yintercept = all_others_data$ET), color = "grey50", linetype="dashed") +
    
    # Add confidence interval band
    geom_rect(aes(ymin = all_others_data$min, ymax = all_others_data$max, xmin = 0, xmax = length(unique(ET_df$age)) + 1), fill = "grey50", alpha = 0.02) +
    
    # Manual legend for the categories and "intermediate age"
    scale_color_manual(name = "", values = c("Years to bearing fruit" = "palegreen2", "Years bearing fruit" = "green3", "Years to removal" = "darkgreen", "Intermediate age" = "grey50")) +
    
    # Axis labels
    xlab("Orchard age") +
    ylab("Agricultural ET (mm/month)") + 
    theme_bw() + 
    theme(legend.position = c(.8, .25), 
        axis.text.x = element_text(angle = 20, hjust = 1)) + 
    ggtitle(crop)
  
  print(plot)
  
  ggsave(here(experiment_path, paste0("Figure_S_", crop, ".pdf")), 
         plot = plot,
         width = 17.8,
         height = 11.1,
         units = "cm")
}
```

